---
title: "Plots - 'artificial VS natural' -experiment"
author: "Ivan Calandra & Lisa Schunk"
date: "`r Sys.time()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---


```{r Knitr Options, include=FALSE}
knitr::opts_chunk$set(comment=NA, message=FALSE, indent="", error=TRUE)

```

---


# Goal of the script
This script plots all variables to see which ones should be used for further analysis.  
Scatterplot of each variable will be plotted.   



```{r}
dir_in <- "analysis/derived_data/"
dir_out <- "analysis/plots"

```

Raw data must be located in `r paste0("~/", dir_in)`.  
Formatted data will be saved in `r paste0("~/", dir_out)`.

The knit directory for this script is the project directory.
---


# Load packages
```{r Libraries}
pack_to_load <- c("tools", "ggplot2", "doBy", "R.utils", "tidyverse", "patchwork", "wesanderson", "ggsci", "ggfortify")
sapply(pack_to_load, library, character.only = TRUE, logical.return = TRUE)

```

---

# Get name, path and information of the file 
```{r}
data_file <- list.files(dir_in, pattern = "\\.Rbin$", full.names = TRUE)
md5_in <- md5sum(data_file)
info_in <- data.frame(file = basename(names(md5_in)), checksum = md5_in, row.names = NULL)

```


# Load data into R object
```{r}
imp_data <- loadObject(data_file)
str(imp_data)

```

The imported file is: "`r paste0("~/", data_file)`"  


---

# Prepare variables
## Define numeric variables
```{r}
num.var <- 22:length(imp_data)

```

The following variables will be used: 

```{r, echo=FALSE}
for (i in num.var) cat("[",i,"] ", names(imp_data)[i], "\n", sep="")

```


---

# Plot each of the selected numeric variable (facet plot = 1 plot for flint, 1 plot for lydite)
```{r}
for (i in num.var){
	#plot
  range_var <- range(imp_data[[i]]) # get the min/max range of the data set 
  
	p_lydite <- ggplot(data = imp_data[grep("LYDIT", imp_data[["Sample"]]), ], 
	                   aes_string(x = "Cycle", y = names(imp_data)[i], 
	                               colour = "Spot")) + 
	            # avoid overplotting
	            geom_point(size = 3) + 
	            geom_line(aes(group = Spot)) + 
	            facet_wrap(Sample+Location ~ Contact.material, nrow = 2) + 
		          coord_cartesian(ylim = range_var) + 
	            ylab(names(imp_data)[i]) + xlab(NULL) +
	            labs(y = gsub("\\.", " ", names(imp_data)[i])) + 
	            scale_colour_futurama() +
	            theme_classic()

	p_flint <- ggplot(data = imp_data[grep("FLT", imp_data[["Sample"]]), ], 
	                  aes_string(x = "Cycle", y = names(imp_data)[i], 
	                             colour = "Spot")) + 
	           geom_point(size = 3) +
	           geom_line(aes(group = Spot)) +
	           facet_wrap(Sample+Location ~ Contact.material, nrow = 2) +
	           coord_cartesian(ylim = range_var) + 
	           ylab(names(imp_data)[i]) + xlab(NULL) +
	           labs(y = gsub("\\.", " ", names(imp_data)[i])) + 
	           scale_colour_futurama() +
	           theme_classic()
	  
	# combine the flint and the lydite plots 
  p <- p_flint + p_lydite + plot_layout(width = c(3/5, 2/5), guides = 'collect')  
       

#save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_plot_", names(imp_data)[i], ".pdf")
	ggsave(filename = file_out, plot = p, path = dir_out, device = "pdf")

}

```


# Plot each of the selected numeric variable in one scatter plot 
```{r}
# exclude the outliers 
# add the indices as row numbers 
imp_data <- imp_data %>% mutate(id = row_number())
imp_data2 <- imp_data[-c(10:12,16:18,19, 20, 58), ]

# add a column that combines sample and location
sample_data <- unite(imp_data2, sample_location, c(Sample, Location), remove = FALSE)
# compute the mean of the three spots per sample 
sample_spot <- summaryBy(.~sample_location+Cycle+Contact.material+Raw.material, 
                         data = sample_data, FUN = mean)
# num.var needs to be defined new since the number of columns changed 
new.num.var <- num.var - 7

sample_spot[["Contact.material"]] <- factor(sample_spot[["Contact.material"]])
# GrandBudapest1 = c("#F1BB7B", "#FD6467", "#5B1A18", "#D67236")
custom.col3 <- data.frame(type = levels(sample_spot$Contact.material), 
                         col = c("#F1BB7B", "#FD6467", "#5B1A18", "#D67236")) 
sample_spot$col <- custom.col3[sample_spot$Contact.material, "col"]


for (i in new.num.var){
	
	p_all <- ggplot(data = sample_spot, aes_string(x = "Cycle", y = names(sample_spot)[i], 
	         colour = "Contact.material", shape = "Raw.material")) +
	         geom_point(size = 3) + 
	         geom_line(aes(group = sample_location)) + 
	         ylab(names(sample_spot)[i]) + xlab(NULL) +
	         labs(y = gsub("\\.", " ", names(sample_spot)[i])) +
	         labs(colour = "Contact material", shape = "Raw material") + 
	         scale_colour_manual(values = custom.col3$col) +  
	         theme_classic()

#save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_VS_plot_", names(sample_spot)[i], ".pdf")
	ggsave(filename = file_out, plot = p_all, path = dir_out, device = "pdf")

}

```


## Principal component analysis 
### PCA contact material
```{r}

# remove rows with na values 
data_pca <- na.omit(imp_data2)
# use for the PCA only selected variables: Sq, SSK, Vmc, Isotropy, Mean density of furrows,
# Asfc, HAsfc9 
imp_data.pca <- prcomp(data_pca[, c(22:23, 40, 45, 49, 52, 54)], scale. = TRUE) 
# convert the data into factor 
data_pca[["Contact.material"]] <- factor(data_pca[["Contact.material"]])


# GrandBudapest1 = c("#F1BB7B", "#FD6467", "#5B1A18", "#D67236")
custom.col3 <- data.frame(type = levels(data_pca$Contact.material), 
                         col = c("#F1BB7B", "#FD6467", "#5B1A18", "#D67236")) 
data_pca$col <- custom.col3[data_pca$Contact.material, "col"]

# Using ggfortify
PCA<- autoplot(imp_data.pca, data = data_pca, colour = "Contact.material", size = 2,
             loadings = TRUE, loadings.colour = "black", loadings.label = TRUE,
             loadings.label.colour = "black", 
             loadings.label.size  = 4, loadings.label.repel = TRUE,  
             #frame = TRUE, frame.type = "convex", frame.colour = "Contact.material",
             frame.alpha = 0) + 
             theme_classic() +
             scale_colour_manual(values = custom.col3$col)
         

# save the plot
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_PCA_contact", ".pdf")
ggsave(filename = file_out, plot = PCA, path = dir_out, device = "pdf")
  
```


### PCA before and after 
```{r}
# convert the data into factor 
data_pca[["Cycle"]] <- factor(data_pca[["Cycle"]])



#  Cavalcanti1 = c("#D8B70A", "#02401B", "#A2A475", "#81A88D", "#972D15")
custom.col6 <- data.frame(type = levels(data_pca$Cycle), 
                         col = c("#D8B70A", "#02401B")) 
data_pca$col <- custom.col6[data_pca$Cycle, "col"]

# using ggfortify
PCA2<- autoplot(imp_data.pca, data = data_pca, colour = "Cycle", size = 2,
             loadings = TRUE, loadings.colour = "black", loadings.label = TRUE, 
             loadings.label.colour = "black", 
             loadings.label.size  = 4, loadings.label.repel = TRUE, frame.alpha = 0) + 
             theme_classic() +
             scale_colour_manual(values = custom.col6$col)
         

# save the plot
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_PCA_cycle", ".pdf")
ggsave(filename = file_out, plot = PCA2, path = dir_out, device = "pdf")
  
```



---

# sessionInfo() and RStudio version

```{r}
sessionInfo()

```

RStudio version `r readLines("analysis/scripts/RStudioVersion.txt", n = 1)`.


# Cite R packages used
```{r Citation}
for (i in pack_to_load) print(citation(i), bibtex = FALSE)

```


---

END OF SCRIPT
