---
title: "Import CSV from ConfoMap ISO25178 - 'artificial VS natural' -experiment"
author: "Lisa Schunk"
date: "`r Sys.time()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---


```{r Knitr Options, include=FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)

```

---


# Goal of the script
This script formats the output of the resulting CSV-file from applying a template computing ISO 25178 parameters in ConfoMap.
The script will:

1. Read in the original CSV-file   
2. Format the data   
3. Write an XLSX-file and save an R object ready for further analysis in R 

```{r}
dir_in <- "analysis/raw_data"
dir_out <- "analysis/derived_data/"

```

Raw data must be located in `r paste0("~/", dir_in)`.  
Formatted data will be saved in `r paste0("~/", dir_out)`.

The knit directory for this script is the project directory.

---

# Load packages
```{r Libraries}
pack_to_load <- c("tools", "openxlsx", "R.utils", "chron")
sapply(pack_to_load, library, character.only = TRUE, logical.return = TRUE)

```


---


# Get names, path and information of the file 
```{r}
data_file <- list.files(dir_in, pattern = "\\.csv$", full.names = TRUE)
md5_in <- md5sum(data_file)
info_in <- data.frame(file = basename(names(md5_in)), checksum = md5_in, row.names = NULL)

```


---


# Read in original CSV-file
```{r}
imp_data <- read.csv(data_file, header = FALSE, stringsAsFactors = FALSE, 
                     na.strings = "*****")
str(imp_data)

```


---


# Format data
## Keep only interesting columns and rows
```{r}
# keep only the columns and rows of interest for the analysis 
data_keep_col <- c(1:2, 16:17, 20:22, 24:26, 29:32, 67:100)
data_keep_rows <- which(imp_data[[1]] != "#")
data_keep <- imp_data[data_keep_rows, data_keep_col]

```


## Add headers
```{r}
head_data_keep <- unlist(imp_data[2, data_keep_col]) 
colnames(data_keep) <- gsub("\\.+", "\\.", make.names(head_data_keep))
colnames(data_keep) <- gsub("\\.$", "", colnames(data_keep)) 

```


## Identify results using frame numbers
```{r}
# combine the results from the different analysis based on the column numbers 
# (ID from MountainsMAp)
frames <- as.numeric(unlist(imp_data[1, data_keep_col]))
ID <- which(frames == 6)[-(1:2)]
ISO <- which(frames == 17)
furrow <- which(frames == 18)
diriso <- which(frames %in% 19:20)
SSFA <- which(frames %in% 21:22)

```


## Shorten the names for parameters
```{r}
# keep only the important information of the headers 
colnames(data_keep)[ISO] <- sapply(strsplit(names(data_keep)[ISO], ".", fixed = TRUE),
                                   `[[`, 1)
colnames(data_keep)[SSFA] <- gsub("^([A-Za-z0-9]+\\.)+", "", colnames(data_keep)[SSFA])

```


## Save units
```{r}
# take the units which were part of the headers and separates them; create a data frame
var_num <- c(ID, ISO, furrow, diriso, SSFA)
# extract 'unit' line for considered columns
units_var <- unlist(imp_data[3, data_keep_col])[var_num] 
# get names associated to the units
names(units_var) <- head_data_keep[var_num] 
# put all of it into a data.frame
units_var_table <- data.frame(variable = names(units_var), unit = units_var) 

```


## Convert to numeric
```{r}
for (i in var_num) data_keep[[i]] <- as.numeric(data_keep[[i]])

```


## Split the column 'Name' into several columns
```{r}
# thes lines extract the artefact ID out of the path name
stud_name <- gsub("^([A-Za-z0-9_]+( --- ))+", "", data_keep[["Name"]])
stud_name <- gsub("([A-Za-z0-9_-]*( - ))+", "", stud_name)
split_name <- do.call(rbind, strsplit(stud_name, "_"))[, 1:4]
split_loc <- do.call(rbind, strsplit(split_name[, 3], "-"))

# split the ID in the separat information 
data_final <- data.frame(split_name[, -3], split_loc, data_keep[-3],
                         stringsAsFactors = FALSE)
colnames(data_final)[1:9] <- c("Sample", "Cycle","Objective", "Location", "Area", "Spot",
                               "Analysis.date", "Analysis.time", "Acquisition.date.time")

# order the the column cycle so that 'before' comes before '2000'
data_final[["Cycle"]] <- factor(data_final[["Cycle"]], levels=c("before", "50", "250", "1000", "2000"))

```


## Format date and time columns
```{r}
data_final[["Analysis.date"]] <- as.Date(data_final[["Analysis.date"]], 
                                         format = "%d.%m.%Y")
data_final[["Analysis.time"]] <- times(data_final[["Analysis.time"]])

```
The column `data_final[["Acquisition.date.time"]]` includes several formats and is therefore
left as `character` without convertion to `POSIXct`.


## Add columns about site, contact material and correct 50x objectives' NAs
```{r}
# extract the raw material based on the ID
data_final[grep("FLT4-", data_final[["Sample"]]), "Raw.material"] <- "flint"
data_final[grep("LYDIT4-", data_final[["Sample"]]), "Raw.material"] <- "lydite"
data_final[["Raw.material"]] <- factor(data_final[["Raw.material"]])

# add the contact/worked material
data_final[grep("LYDIT4-1", data_final[["Sample"]]), "Contact.material"] <- "pork skin"
data_final[grep("LYDIT4-2", data_final[["Sample"]]), "Contact.material"] <- "bone plate"
data_final[grep("LYDIT4-5", data_final[["Sample"]]), "Contact.material"] <- "bos scapula"
data_final[grep("LYDIT4-9", data_final[["Sample"]]), "Contact.material"] <- "skin pad"

data_final[grep("FLT4-4", data_final[["Sample"]]), "Contact.material"] <- "pork skin"
data_final[grep("FLT4-5", data_final[["Sample"]]), "Contact.material"] <- "bone plate"
data_final[grep("FLT4-15", data_final[["Sample"]]), "Contact.material"] <- "bos scapula"
data_final[grep("FLT4-12", data_final[["Sample"]]), "Contact.material"] <- "skin pad"

# correct information about the used objectives 
data_final[data_final[["Objective"]] == "50x09", "Objective"] <- "50x095"
data_final[data_final[["Objective"]] == "50x07", "Objective"] <- "50x075"
data_final[["Objective"]] <- factor(data_final[["Objective"]])

```


## Ignore some columns and reorder columns
```{r}
data_final <- data_final[c(1:2, 4:6, 3, 54:55, 7:9, 10:53)]

```


## Add units as comment()
```{r}
comment(data_final) <- units_var

```

Type `comment(data_final)` to check the units of the columns.


## Check the result
```{r}
str(data_final)
head(data_final)

```


---


# Save data
## Format name of output file
```{r}
file_out <- "AvsN_use-wear"

```
The files will be saved as "`r paste0("~/", file_out, ".[ext]")`".


## Write to XLSX
```{r}
write.xlsx(list(data = data_final, units = units_var_table), 
           file = paste0(dir_out, file_out, ".xlsx"))

```


## Save R object
```{r}
saveObject(data_final, file = paste0(dir_out, file_out, ".Rbin"))

```

---

# sessionInfo() and RStudio version

```{r}
sessionInfo()
```

RStudio version `r readLines("analysis/scripts/RStudioVersion.txt", n = 1)`.


# Cite R packages used
```{r Citation}
for (i in pack_to_load) print(citation(i), bibtex = FALSE)

```

---


END OF SCRIPT
